---
title: "Bootstrapping"
author: "Yuxuan Wang"
date: "2023-11-16"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

# Generate a relevant example

```{r}
n_samp = 250

sim_df_const = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )

sim_df_nonconst = sim_df_const |> 
  mutate(
  error = error * .75 * x,
  y = 2 + 3 * x + error
)
```

```{r}
sim_df = 
  bind_rows(const = sim_df_const, nonconst = sim_df_nonconst, .id = "data_source") 

sim_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm") +
  facet_grid(~data_source) 
```

```{r}
lm(y ~ x, data = sim_df_const) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)

sim_df_nonconst |> 
  lm(y ~ x, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```


# Drawing one bootstrap sample

Start with a little function

```{r}
boot_sample = function(df){
  
  sample_frac(df, replace = TRUE)
}
```

Let's see how it works

```{r}
sim_df_nonconst |> 
  boot_sample() |> 
  ggplot(aes(x = x, y = y)) +
  geom_point(alpha = .5) +
  stat_smooth(menthos = "lm")
```

# Drawing many bootstrap samples


```{r}
boot_straps = 
  tibble(strap_number = 1:1000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = sim_df_nonconst))
  )

boot_straps
```

```{r}
boot_straps |> 
  slice(1:3) |> 
  mutate(strap_sample = map(strap_sample, arrange, x)) |> 
  pull(strap_sample)
```

```{r}
boot_straps |> 
  slice(1:3) |> 
  unnest(strap_sample) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm", se = FALSE) +
  facet_grid(~strap_number) 
```

# Analyzing bootstrap samples

Now do the `lm` fit.

```{r}
bootstrap_results = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(y ~ x, data = df) ),
    results = map(models, broom::tidy)) |> 
  select(-strap_sample, -models) |> 
  unnest(results) 
```

Try to summarize these results -- get a bootstrap SE

```{r}
bootstrap_results |> 
  group_by(term) |> 
  summarize(boot_se = sd(estimate)) |> 
  knitr::kable(digits = 3)
```

Look at the distrubution

```{r}
bootstrap_results |> 
  filter(term == "x") |> 
  ggplot(aes(x = estimate)) +
  geom_density()
```

Can I construct a CI

```{r}
bootstrap_results |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975))
```

```{r}
boot_straps |> 
  unnest(strap_sample) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_line(aes(group = strap_number), stat = "smooth", method = "lm", se = FALSE, alpha = .1, color = "blue") +
  geom_point(data = sim_df_nonconst, alpha = .5)
```

# Airbnb data

Let's fit a regression of `price` on other variables and look at residuals

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  drop_na(price, stars) |> 
  select(price, stars, borough, neighborhood, room_type)

nyc_airbnb |> 
  ggplot(aes(x = stars, y = price, color = room_type)) + 
  geom_point()
```


run a bootstrap in this whole thing ti get estimate for the effect of `stars` on `price`

```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  modelr::bootstrap(n = 1000) |> 
  mutate(
    models = map(strap, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(results) |> 
  unnest(results) |> 
  filter(term == "stars") |> 
  ggplot(aes(x = estimate)) + geom_density()
```

