---
title: "linear_models"
author: "Yuxuan Wang"
date: "2023-11-11"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

# Model fitting

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

# fit

```{r}
summary(fit)
summary(fit)$coef
coef(fit)
```

# Tidying output

```{r}
fit |> 
  broom::glance()
```

# Tidy the coefficients

```{r}
fit |> 
  broom::tidy()
```

```{r}
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  knitr::kable(digits = 3)
```

# Diagnostics

```{r}
modelr::add_residuals(nyc_airbnb, fit)

modelr::add_predictions(nyc_airbnb, fit)
```

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = resid)) + 
  geom_density() +
  xlim(-100,500)
```

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = borough, y = resid)) + geom_violin()
```

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = stars, y = resid)) + geom_point()
```

# Hypothesis testing

Fit a "null" and "alternative" model 

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

```{r}
anova(fit_null, fit_alt) |> 
  broom::tidy()
```

# Nesting data

```{r}
nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

```{r}
nest_lm_res =
  nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    models = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(-data, -models) |> 
  unnest(results)
```

```{r}
nest_lm_res |> 
  select(borough, term, estimate) |> 
  mutate(term = fct_inorder(term)) |> 
  pivot_wider(
    names_from = term, values_from = estimate) |> 
  knitr::kable(digits = 3)
```

```{r}
manhattan_airbnb =
  nyc_airbnb |> 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb |> 
  nest(data = -neighborhood) |> 
  mutate(
    models = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(-data, -models) |> 
  unnest(results)
```

```{r}
manhattan_nest_lm_res |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

```{r}
manhattan_airbnb |> 
  lme4::lmer(price ~ stars + room_type + (1 + room_type | neighborhood), data = _) |> 
  broom.mixed::tidy()
```

# Binary outcomes

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 
```

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value) |> 
  knitr::kable(digits = 3)
```

```{r}
baltimore_df |> 
  modelr::add_predictions(fit_logistic) |> 
  mutate(fitted_prob = boot::inv.logit(pred))
```


## Nolinear data and CV

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point()
```

Do the train / test split.

```{r}
train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```

quick visualization of the linear model


```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

```{r}
train_df |> 
  modelr::add_predictions(linear_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

```{r}
train_df |> 
  modelr::add_predictions(smooth_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

```{r}
train_df |> 
  modelr::add_predictions(wiggly_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

RMSEs on tresting data gives a sense if out-of-sample prediction accurancy!
```{r}
rmse(linear_mod, test_df)

rmse(smooth_mod, test_df)

rmse(wiggly_mod, test_df)
```


# CV using modelr

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) 

cv_df |> pull(train) |> nth(1) |> as_tibble()
```

```{r}
cv_df |> pull(test) |> nth(1) |> as_tibble()
```

Apply each model to all training datasets, and evaluate an all testing datasets

```{r}
cv_df =
  cv_df |> 
  mutate(
    linear_fit = map(train, \(df) lm(y ~ x, data = df))) |> 
  mutate(
    rmse_linear = map2(linear_fit, test, \(mod, df ) rmse(mod, df))
  )
```

```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, ~lm(y ~ x, data = .x)),
    smooth_mod  = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
    wiggly_mod  = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))
    ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y))
    )
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```



